---
title: "Seattle Car Crashes"
author: "Simoli Aghara, Mindy Chen, Rachel Graham, Tammy Le"
date: "`r Sys.Date()`"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
```


About
===================================== 

### About Section

This is about Seattle Car Crashes. 

Table
===================================== 

```{r}
library(readxl)
library(tidyverse)
library(DT)
Seattle_Car_Crashes <- read_excel("Seattle Car Crashes.xlsx")
Seattle_Car_Crashes <- na.omit(Seattle_Car_Crashes)
Seattle_Car_Crashes <- subset(Seattle_Car_Crashes, select = -c(INCDTTM))
datatable(Seattle_Car_Crashes)

```


Geolocation
===================================== 

Scatter + Line
===================================== 

Scatter plot: Correlation between weather conditions, daylight vs nighttime, and road conditions. - Rachel 
Line Graph: Correlation between date, and time(daylight vs nighttime) - Simoli 


Scatter + Bar
===================================== 

Bar Graph: Type of Accident and Junction and Severity - Mindy
Scatter plot: Compare most injuries/severies to Junction (intersections) - Tammy 
