---
title: "Seattle Car Crashes"
author: "Simoli Aghara, Mindy Chen, Rachel Graham, Tammy Le"
date: "`r Sys.Date()`"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(readxl)
library(tidyverse)
library(DT)
library(dplyr)
library(ggmap)
library(ggplot2)
library(plotly)
library(shiny)
library(lubridate)
library(leaflet)
```


About
===================================== 

### About Section

##### Introduction:
Seattle's frequent car accidents, often attributed to its grey weather and rainy conditions, prompt the need for a comprehensive understanding of these incidents. This dashboard aims to shed light on car crashes in Seattle from January 1, 2018, to February 8, 2024, using a series of visualizations.

##### Data Table Overview:
The dashboard begins with a concise overview of the dataset in tabular form, providing users with a foundational understanding of the data used to generate subsequent visualizations.

##### Temporal Analysis Map:
A dynamic map illustrates the spatial distribution of car crashes in Seattle over time. Users can observe patterns, hotspots, and temporal trends to gain insights into accident occurrences across different neighborhoods and roadways.

##### Attribution of Accident Causes:
Delving deeper, the dashboard analyzes the primary causes of accidents—be it adverse weather conditions, road infrastructure, or inadequate lighting. The visualization has been divided into two parts to provide better understanding. These visualization elucidates the factors contributing most significantly to car crashes, aiding in pinpointing areas for potential safety enhancements.

##### Injury Severity Analysis:
Furthermore, the dashboard explores the types of accidents that result in the most severe injuries, identifying both the location and nature of incidents. This critical insight enables stakeholders to prioritize interventions aimed at reducing the incidence of severe injuries in specific contexts.

##### Location-Based Analysis:
A breakdown of accident-prone locations reveals where the majority of crashes occur, offering valuable insights into high-risk areas. Notably, the visualization highlights common scenarios such as left-turn accidents, providing actionable information for targeted intervention strategies.

##### Limitations:
Acknowledging the limitations, it's important to note the absence of data for weather, road, and light conditions in the year 2024. This constraint impacts the comprehensiveness of the analysis for the most recent period.

##### Objective and Impact:
In essence, the dashboard seeks to address the challenge of designing an accessible platform that synthesizes essential information about car collisions in Seattle. By facilitating a deeper understanding of accident attributes, the dashboard aims to empower auto engineers to implement safety adjustments and policymakers to enact targeted traffic safety measures. Ultimately, the overarching goal is to raise public awareness of the primary factors contributing to car crashes and foster a safer urban environment for all.

##### GitHub Link:

Table
===================================== 

```{r Reading and Displaying Data}
Seattle_Car_Crashes <- read_excel("Seattle_Car_Crashes_2018-2024.xlsx")
Seattle_Car_Crashes <- na.omit(Seattle_Car_Crashes)
Seattle_Car_Crashes <- subset(Seattle_Car_Crashes, select = -c(INCDTTM))
datatable(Seattle_Car_Crashes)

```

Geolocation Map
===================================== 

```{r GeoLocation}
location <- Seattle_Car_Crashes %>%
  mutate(year = year(INCDATE), month = month(INCDATE)) %>% 
  select(LATITUDE,LONGITUDE,LOCATION,INCDATE,COLLISIONTYPE,JUNCTIONTYPE,
         WEATHER,ROADCOND,LIGHTCOND,SEVERITYCODE,year)

ui <- fluidPage(
  titlePanel("Accident Map"),
  sidebarLayout(
    mainPanel(
      selectInput("year", "Select Year:", choices = unique(location$year))
    ),
    mainPanel(
      leafletOutput("map", width = "800px", height = "600px")
    )
  )
)

# Define server logic
server <- function(input, output) {
  
  # Filter accidents based on selected year
  data_2 <- reactive({
    subset(location, year == input$year)
  })
  
  pal <- colorFactor(
    palette = c("lightyellow", "yellow", "orange", "red", "darkred"),
    domain = c("0", "1", "2", "2b", "3")
  )
  
  # Render leaflet map
  output$map <- renderLeaflet({
    leaflet(data = data_2()) %>%
      addTiles() %>%
      addCircleMarkers(
        lng = ~LONGITUDE,
        lat = ~LATITUDE,
        radius = 5,
        label = ~paste("<span style='font-size: 13px;'>", 
                       "<strong>", format(as.Date(INCDATE), "%Y-%m-%d"), "</strong>", "</span>",
                       "<strong><br>Location:</strong>", LOCATION,
                       "<strong><br>Collision Type:</strong>", COLLISIONTYPE,
                       "<strong><br>Junction Type:</strong>", JUNCTIONTYPE,
                       "<strong><br>Weather:</strong>", WEATHER,
                       "<strong><br>Road Condition:</strong>", ROADCOND,
                       "<strong><br>Light Condition:</strong>", LIGHTCOND,
                       "<strong><br>Severity Code:</strong>", SEVERITYCODE) %>%
                        lapply(htmltools::HTML),
        color = ~pal(SEVERITYCODE),
        opacity = 0.7,
        fillOpacity = 0.7
      ) %>%
      # Define color palette
      addLegend(position = "bottomright", 
                pal = pal,
                values = ~SEVERITYCODE,
                title = "Severity Code"
      )
  })
}

# Run the application
shinyApp(ui = ui, server = server)

```


Road and Weather Condition
===================================== 

```{r Scatter Plot}
Seattle_Car_Crashes_Date <- Seattle_Car_Crashes %>%
  mutate(year = year(INCDATE), month = month(INCDATE))

total_accident <- Seattle_Car_Crashes_Date %>%
  group_by(year, month, WEATHER, ROADCOND) %>%
  summarise(Number_Accidents = n())

ui <- fluidPage(
  titlePanel("Seattle Car Crashes: Weather and Road Conditions by Year"),
  
  # Adding Information about the visualization
  fluidRow(
    column(
      width = 12,
      h3("Visualization Description"),
      p("The following scatterplots present a detailed analysis of how weather and road conditions have influenced car accidents over time in Seattle. It aims to determine whether certain combinations of weather and road conditions contribute to accident occurrences. Users have the flexibility to select the desired year for analysis, enabling a focused examination of trends and patterns."),
      br(),  # Add a line break for spacing
    )
  ),
  
  fluidRow(
    column(
          width = 8,
          selectInput("year","Select Year",choices = unique(total_accident$year))
      )
  ),

  fluidRow(
    column(
          width = 12,
          plotlyOutput("plot", width = "90%", height = "600px")
    )
  )
)

server <- function(input, output) {
  Seattle_Car_Crashes_Date <- Seattle_Car_Crashes %>%
    mutate(year = year(INCDATE), month = month(INCDATE))
  # Data processing
  Seattle_Car_Crashes <- Seattle_Car_Crashes %>%
    mutate(Weather_Combined = case_when(
      WEATHER %in% c("Clear", "Overcast", "Partly Cloudy") ~ "Clear",
      WEATHER %in% c("Fog/Smog/Smoke") ~ "Fog",
      WEATHER %in% c("Raining", "Sleet/Hail/Freezing Rain") ~ "Rain",
      WEATHER %in% c("Snowing", "Blowing Snow") ~ "Snow",
      WEATHER %in% c("Other", "Blowing Sand/Dirt", "Severe Crosswind") ~ "Other",
      TRUE ~ WEATHER  # Keep other values unchanged
    )) %>%
    mutate(Road_Combined = case_when(
      ROADCOND == "Dry" ~ "Dry",
      ROADCOND %in% c("Wet", "Standing Water", "Snow/Slush", "Sand/Mud/Dirt", "Oil") ~ "Wet",
      ROADCOND == c("Ice") ~ "Ice",
      TRUE ~ ROADCOND  # Keep other values unchanged
    )) %>%
    filter(WEATHER != "Unknown" & ROADCOND != "Unknown" & ROADCOND != "Other")


  # Output the interactive plot
  output$plot <- renderPlotly({
    
    # Filter data based on user-selected year
    filtered_data <- Seattle_Car_Crashes %>%
      mutate(year = year(INCDATE), month = month(INCDATE)) %>% 
      filter(year == input$year) %>%
      group_by(month, WEATHER, ROADCOND, INCDATE, Road_Combined, Weather_Combined) %>%
      summarise(Number_Accidents = n())
    
    p <- ggplot(filtered_data, aes(x = month, y = Number_Accidents, color = Road_Combined, shape = Weather_Combined, text = paste("Weather:", Weather_Combined, "<br>Road:", Road_Combined, "<br>Crashes:", Number_Accidents))) +
      geom_point(size = 3) +
      scale_x_discrete(limits = 1:12, labels = month.abb) +
      labs(x = "Year-Month", y = "Number of Car Crashes", title = paste("Car Crashes by Weather and Road Condition in", input$year), color = "Road Condition", shape = "Weather Condition") +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "top", panel.background = element_rect(fill = "lightgrey"))
    
    ggplotly(p) %>%
layout(legend = list(orientation = "h"))
    
    ggplotly(p, tooltip = "text")
  })
}

shinyApp(ui = ui, server = server)
```

Light Condition
===================================== 

```{r Line Graph}
# Separating the date column from the dataset
Seattle_Car_Crashes_Date <- Seattle_Car_Crashes %>% select(INCDATE, LIGHTCOND) %>% drop_na() %>%
  mutate(year = year(INCDATE), month = month(INCDATE))

# Counting the total number of accidents happened in a particular month
total_accident <- Seattle_Car_Crashes_Date %>%
  group_by(LIGHTCOND, year, month) %>%
  summarise(Number_Accidents = n())

# Making the graph interactive
ui <- fluidPage(
  
  # Giving the title to the graph
  titlePanel("Car Crashes by Light Condition over Time"),
  
  # Adding Information about the visualization
  fluidRow(
    column(
      width = 12,
      h3("Visualization Description"),
      p("The following visualization presents a detailed analysis of how light conditions have influenced car accidents over time in Seattle. It aims to elucidate whether insufficient lighting contributes to accident occurrences. Users have the flexibility to select the desired year for analysis, enabling a focused examination of trends and patterns."),
      br(),  # Add a line break for spacing
    )
  ),
  
  # Lets the user choose which year they want to view
  sidebarLayout(
    mainPanel(
      fluidRow(
        column(
          width = 8,
          selectInput("year","Select Year",choices = unique(total_accident$year))
        )
      )
    ),
    
    # The output
    mainPanel(
      fluidRow(
        column(
          width = 12,
          plotlyOutput("plot", width = "150%", height = "600px")
        )
      )
    )
  )
)

server <- function(input, output) {
  # Creating the data with the year selected by the user
  data <- reactive({
    filter(total_accident, year == input$year)
  })
  
  output$plot <- renderPlotly({
     lightcond_colors <- c("Daylight" = "blue", "Dark - Street Lights On" = "green", "Dusk" = "orange", "Dawn" = "red", "Dark - No Street Lights" = "purple", "Dark - Street Lights Off" = "brown", "Unknown" = "gray", "Other" = "black")
    # Creating the plot
    p <- ggplot(data(), aes(x = month, y = Number_Accidents, group = LIGHTCOND, color = LIGHTCOND, text = paste("Crashes:", Number_Accidents))) +
      geom_line() +
      scale_x_discrete(limits = 1:12, labels = month.abb) +
      labs(x = "Year-Month", y = "Number of Accidents", title = paste("Car Crashes by Light Condition in", input$year)) +
       scale_color_manual(values = lightcond_colors) + 
      theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels by 45 degrees
    
     ggplotly(p) %>%
layout(legend = list(orientation = "h"))
    
    ggplotly(p, tooltip = "text")

  })
}

shinyApp(ui = ui, server = server)
```

Severity of injuries vs Junction type
===================================== 

```{r Scatter_Plot2}
# Read the dataset from Excel file
Seattle_Car_Crashes <- read_excel("Seattle_Car_Crashes_2018-2024.xlsx")

# Filter out rows with unknown or irrelevant junction types
valid_junction_types <- c("At Intersection (but not related to intersection)",
                          "At Intersection (intersection related)",
                          "Driveway Junction",
                          "Mid-Block (but intersection related)",
                          "Mid-Block (not related to intersection)")
Seattle_Car_Crashes <- Seattle_Car_Crashes %>%
  filter(JUNCTIONTYPE %in% valid_junction_types) %>%
  mutate(year = year(INCDATE))  # Add a year column for filtering

# UI
ui <- fluidPage(
  titlePanel("Seattle Car Crashes: Number of Injuries and Severity Level at Junction Types"),
  
  # Visualization Description
  fluidRow(
    column(
      width = 12,
      h3("Visualization Description"),
      p("This visualization provides insights into the correlation between the number of injuries and the severity level of car crashes at different junction types in Seattle. It aims to highlight the impact of junction type on the severity and outcomes of accidents. Users can explore the data to understand which junction types are most associated with severe accidents."),
      br()  # Add a line break for spacing
    )
  ),
  
  # Year selection above the graph
  fluidRow(
    column(
      width = 12,
      selectInput("year", "Select Year:", choices = unique(Seattle_Car_Crashes$year))
    )
  ),
  
  # Graph
  fluidRow(
    column(
      width = 12,
      plotlyOutput("crash_plot", width = "100%", height = "600px")
    )
  )
)

# Server logic
server <- function(input, output) {
  # Reactive expression to filter data based on selected year
  filtered_data <- reactive({
    data <- Seattle_Car_Crashes %>% filter(year == input$year)
    
    # Calculate total injuries per JUNCTIONTYPE
    injuries_per_junction <- data %>%
      group_by(JUNCTIONTYPE) %>%
      summarise(Total_Injuries = sum(INJURIES)) %>%
      arrange(desc(Total_Injuries)) # Arrange by descending order of total injuries

    # Reorder the JUNCTIONTYPE factor based on the total injuries
    data$JUNCTIONTYPE <- factor(data$JUNCTIONTYPE, levels = injuries_per_junction$JUNCTIONTYPE)
    
    data
  })
  
  # Output the interactive plot with faceting and without the legend
  output$crash_plot <- renderPlotly({
    # Create a faceted plot without a legend
    p <- ggplot(filtered_data(), aes(x = SEVERITYCODE, y = INJURIES, color = JUNCTIONTYPE, shape = JUNCTIONTYPE, text = paste("Severity Code:", SEVERITYCODE, "<br>Number of Injuries:", INJURIES, "<br>Junction Type:", JUNCTIONTYPE))) +
      geom_jitter(size = 4, alpha = 0.5) + # Apply jittering and transparency
      facet_wrap(~JUNCTIONTYPE, scales = "free_y") + # Faceting by junction type
      scale_shape_manual(values = c(16, 17, 15, 4, 2)) + # Different shapes for junction types
      labs(x = "Severity Level", y = "Number of Injuries") +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1), 
            panel.background = element_rect(fill = "lightgrey"), 
            strip.background = element_rect(fill = "lightblue"), 
            strip.text = element_text(size = 10),
            legend.position = "none") # Remove the legend
    
    # Convert ggplot object to plotly object to add tooltips
    ggplotly(p, tooltip = "text")
  })
}

# Run the application
shinyApp(ui = ui, server = server)
```

Collision type VS Junction type
===================================== 

```{r}
library(plotly)
library(dplyr)

# Calculate counts of each junction type and severity level
junction_severity_counts <- Seattle_Car_Crashes %>%
  group_by(JUNCTIONTYPE, SEVERITYCODE) %>%
  summarise(count = n()) %>%
  ungroup()

# Create a plotly bar chart with click events
plot_ly(junction_severity_counts, x = ~JUNCTIONTYPE, y = ~count, type = 'bar', color = ~SEVERITYCODE, source = 'data') %>%
  layout(
    title = 'Relationship Between Junction Type and Severity',
    xaxis = list(title = 'Junction Type', categoryorder = 'total descending'),
    yaxis = list(title = 'Count'),
    barmode = 'stack'
  ) %>%
  highlight(selectize = TRUE) %>%
  event_register('plotly_click')


```

